\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-12-02 paper.qmd
Added initial set up for UMAP with PBMC..our method is working !!!!

>>>>

#### scDEED

**tSNE**

```{r}
result_tsne <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_tsne_results.rds'))
```

```{r}
result_tsne |>
  kableExtra::kable(format = "latex",
                    booktabs = TRUE,
                    longtable = TRUE,
                    label = "scdeedtsne") |>
  kableExtra::kable_styling(latex_options = "scale_down")
```

```{r}
#| out-width: 50%

ggplot(data = result_tsne,
       aes(
         x = as.numeric(perplexity),
         y = number_dubious_cells
       )) +
  geom_line() +
  geom_point() +
  labs(x = "perplexity",
       y = "# of dubious cells") +
  theme_minimal()
```

**UMAP**

```{r}
result_umap <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_umap_results.rds'))
```

```{r}
result_umap |>
  kableExtra::kable(format = "latex",
                    booktabs = TRUE,
                    longtable = TRUE,
                    label = "scdeedumap") |>
  kableExtra::kable_styling(latex_options = "scale_down")
```

```{r}
#| out-width: 50%

result_umap |>
  filter(min.dist == 0.5) |>
  ggplot(aes(
           x = as.numeric(n_neighbors),
           y = number_dubious_cells
         )) +
    geom_line() +
    geom_point() +
    labs(x = "n_neighbors",
         y = "# of dubious cells") +
    theme_minimal()
```

```{r}
#| label: read-pbmc-umap-layouts
# Read a variety of different NLDR representations of PBMC
# and plot them on same aspect ratio
umap_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds")
umap_pbmc1 <- as_tibble(umap_pbmc1)
names(umap_pbmc1) <- c("UMAP1", "UMAP2")

nldr1c <- umap_pbmc1 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("a")

umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
umap_pbmc2 <- as_tibble(umap_pbmc2)
names(umap_pbmc2) <- c("UMAP1", "UMAP2")

nldr2c <- umap_pbmc2 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#377eb8') +
  interior_annotation("b")

```{r}
#| label: combine-data-pbmc-umap

error_pbmc_umap <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.3.rds")
error_pbmc_umap2 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_80_min_dist_0.5.rds")

error_pbmc <- bind_rows(error_pbmc_umap, 
                        error_pbmc_umap2)

```{r}
#| label: error-comp-pbmc-umap

error_umap_plot_pbmc <- ggplot(error_pbmc, 
                               aes(x = a1, 
                                   y = MSE, 
                                   colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
  scale_color_manual(values=c('#e41a1c','#377eb8')) +
  scale_y_log10() +
  ylab("log(MSE)") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

```{r}
#| label: fig-pbmc-mse-umap
#| fig-pos: H
#| out-height: 100%

free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2)
```

### 2024-12-02 paper.tex
Added initial table for scDEED
Added 3 layouts for three UMAP hyperparameters
Added results for different UMAP parameter choices
Set the seed to get the same results as before
Added initial set up for UMAP with PBMC..our method is working !!!!

>>>>

\subsubsection{scDEED}\label{scdeed}

\textbf{tSNE}

\begin{longtable*}{lr}
\toprule
perplexity & number\_dubious\_cells\\
\midrule
5 & 87\\
10 & 143\\
15 & 146\\
20 & 101\\
25 & 93\\
\addlinespace
30 & 83\\
35 & 82\\
40 & 85\\
45 & 71\\
50 & 41\\
\addlinespace
55 & 48\\
60 & 46\\
65 & 35\\
70 & 36\\
75 & 49\\
\addlinespace
80 & 46\\
85 & 75\\
90 & 82\\
95 & 62\\
100 & 20\\
\addlinespace
105 & 19\\
110 & 19\\
115 & 21\\
120 & 18\\
125 & 17\\
\addlinespace
130 & 15\\
135 & 17\\
140 & 17\\
145 & 18\\
\bottomrule
\end{longtable*}

\begin{center}
\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-62-1.pdf}
\end{center}

\textbf{UMAP}

\begin{longtable*}{rrr}
\toprule
n\_neighbors & min.dist & number\_dubious\_cells\\
\midrule
5 & 0.0125 & 14\\
6 & 0.0125 & 7\\
7 & 0.0125 & 5\\
8 & 0.0125 & 10\\
9 & 0.0125 & 29\\
\addlinespace
10 & 0.0125 & 7\\
20 & 0.0125 & 9\\
30 & 0.0125 & 10\\
40 & 0.0125 & 4\\
50 & 0.0125 & 3\\
\addlinespace
80 & 0.0125 & 47\\
160 & 0.0125 & 13\\
240 & 0.0125 & 74\\
5 & 0.0500 & 12\\
6 & 0.0500 & 37\\
\addlinespace
7 & 0.0500 & 6\\
8 & 0.0500 & 15\\
9 & 0.0500 & 33\\
10 & 0.0500 & 5\\
20 & 0.0500 & 13\\
\addlinespace
30 & 0.0500 & 4\\
40 & 0.0500 & 4\\
50 & 0.0500 & 8\\
80 & 0.0500 & 15\\
160 & 0.0500 & 55\\
\addlinespace
240 & 0.0500 & 75\\
5 & 0.1000 & 271\\
6 & 0.1000 & 9\\
7 & 0.1000 & 16\\
8 & 0.1000 & 37\\
\addlinespace
9 & 0.1000 & 4\\
10 & 0.1000 & 20\\
20 & 0.1000 & 5\\
30 & 0.1000 & 5\\
40 & 0.1000 & 5\\
\addlinespace
50 & 0.1000 & 25\\
80 & 0.1000 & 4\\
160 & 0.1000 & 4\\
240 & 0.1000 & 13\\
5 & 0.3000 & 45\\
\addlinespace
6 & 0.3000 & 10\\
7 & 0.3000 & 12\\
8 & 0.3000 & 4\\
9 & 0.3000 & 3\\
10 & 0.3000 & 3\\
\addlinespace
20 & 0.3000 & 2\\
30 & 0.3000 & 46\\
40 & 0.3000 & 2\\
50 & 0.3000 & 3\\
80 & 0.3000 & 2\\
\addlinespace
160 & 0.3000 & 2\\
240 & 0.3000 & 21\\
5 & 0.5000 & 88\\
6 & 0.5000 & 3\\
7 & 0.5000 & 5\\
\addlinespace
8 & 0.5000 & 4\\
9 & 0.5000 & 92\\
10 & 0.5000 & 5\\
20 & 0.5000 & 72\\
30 & 0.5000 & 70\\
\addlinespace
40 & 0.5000 & 1\\
50 & 0.5000 & 2\\
80 & 0.5000 & 2\\
160 & 0.5000 & 3\\
240 & 0.5000 & 4\\
\addlinespace
5 & 0.7000 & 13\\
6 & 0.7000 & 86\\
7 & 0.7000 & 95\\
8 & 0.7000 & 29\\
9 & 0.7000 & 95\\
\addlinespace
10 & 0.7000 & 6\\
20 & 0.7000 & 3\\
30 & 0.7000 & 5\\
40 & 0.7000 & 2\\
50 & 0.7000 & 3\\
\addlinespace
80 & 0.7000 & 4\\
160 & 0.7000 & 5\\
240 & 0.7000 & 4\\
5 & 0.9000 & 192\\
6 & 0.9000 & 95\\
\addlinespace
7 & 0.9000 & 233\\
8 & 0.9000 & 228\\
9 & 0.9000 & 101\\
10 & 0.9000 & 7\\
20 & 0.9000 & 3\\
\addlinespace
30 & 0.9000 & 79\\
40 & 0.9000 & 3\\
50 & 0.9000 & 5\\
80 & 0.9000 & 4\\
160 & 0.9000 & 5\\
\addlinespace
240 & 0.9000 & 3\\
\bottomrule
\end{longtable*}

\begin{center}
\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-65-1.pdf}
\end{center}

\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}

\caption{\label{fig-pbmc-mse-umap}}
